---
title: "manuscript_update"
author: "Binchen Mao"
date: "2023-08-18"
output:
    html_document:
        code_folding: hide
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = F,
  warning = F,
  fig.height = 8,
  fig.width = 16,
  dpi = 72,
  cache = FALSE
)
```

```{r setup}
library(invivoSyn)
library(ggplot2)
library(ggpubr)
library(pracma)
library(magrittr)
library(kableExtra)
library(dplyr)

simu_TV2=function(n=5,tv0_min=100,tv0_max=300,vehicle_double_time=3,t=21,cv_r=0.1,eu_tv=3000,tgi_a=0.5,tgi_b=0.4,tgi_c=0.8,kit_a=0,kit_b=0,kit_c=0,rp_a=0,rp_b=0,rp_c=0,irt_a=0,irt_b=0,irt_c=0,sigma=1,...){
  trs=c('Vehicle control (C)','Drug A (A)','Drug B (B)','Drug A+B (A+B)')
  treatments=factor(trs,levels=trs)
  tv0s=as.vector((sapply(1:4,function(i) runif(n,min=tv0_min,max=tv0_max))))
  k_v=log(2)/vehicle_double_time #calcul-1ate growth rate for vehicle group
  sd_k=cv_r*k_v #standard deviation of tumor growth rate
  drug_eff_vec=rep(c(0,log(1-tgi_a)/14,log(1-tgi_b)/14,log(1-tgi_c)/14),each=n)#drug effect
  kit_vec=rep(c(0,kit_a,kit_b,kit_c),each=n) #kickin time vector
  rp_vec=rep(c(0,rp_a,rp_b,rp_c),each=n)
  irt_vec=rep(c(0,irt_a,irt_b,irt_c),each=n) #induced resistance time vector
  k_vec=as.vector(sapply(1:4,function(i) rnorm(n,mean=k_v,sd=sd_k)))#vector of initial tumor growth rate
  #k_vec=as.vector(sapply(c(k_v,k_a,k_b,k_c),function(k) rnorm(n,mean=k,sd=sigma_k)))
  tv=data.frame(Treatment=rep(treatments,each=n),Mouse=1:(4*n),Day=0,TV=tv0s,kit=kit_vec,rp=rp_vec,irt=irt_vec,k_c=k_vec,drug_eff=drug_eff_vec) %>%
    mutate(k=ifelse(kit==0,k_c+drug_eff,k_c),rpv=ifelse(irt==0,rp_vec*TV,0))
  tvs=list()
  tvs[[1]]=tv
  for(i in 1:t){
    tvs[[i+1]]=tvs[[i]] %>% mutate(Day=i) %>% mutate(k=ifelse(Day>=kit,k_c+drug_eff,k_c),rpv=ifelse(Day>=irt & rpv==0,rp*TV,rpv)) %>% mutate(TV=(TV-rpv)*exp(k)+rpv*exp(k_c),rpv=rpv*exp(k_c)) %>%
      mutate(TV=TV+rnorm(n*4,sd=sigma)) %>% mutate(TV=ifelse(TV>=eu_tv,NA,TV))
  }
  tv_long=do.call(rbind,tvs) %>% as.data.frame()
  tv_long=tv_long %>% mutate(Group=factor(paste0("Group ",as.numeric(Treatment)))) %>% select(Group,Treatment,Mouse,Day,TV)
  tv_long=expand_tv(tv_long)
  tv_long
}
```

# Introduction

This rmarkdown file includes the code to generate the figures used in manuscript **Statistical Assessment of Drug Synergy from In Vivo Combination Studies using Mouse Tumor Models**.

simulate tumor growth date to generate Figure 1B, then calculate average eGR of each group to estimate tumor growth rate under exponential tumor growth model, and plot tumor growth curve as Figure 1C. Finally, use invivoSyn's method to evaluate Bliss synergy (Figure 1D) 

# Figure 1

```{r Figure1}
#--draw Figure 1A
#--sample from exponential growth--#
#set.seed(12345)
set.seed(123451)
day = c(0, 3, 7, 10, 14, 17, 21, 24,27,30)
V0 = 50
k = 0.11
Vt = V0 * exp(k*day)
n = length(Vt)
Vtt = Vt + rnorm(n, Vt, 5*day)
TVday = data.frame(day,log(Vtt), log(Vtt[1]))
names(TVday) = c('day','logTV','logTV0')
TVday.AUC = pracma::trapz(x=TVday$day,y=TVday$logTV) - pracma::trapz(x=TVday$day,y=TVday$logTV0)
TVday.eg = function(x) log(100) + 0.1077018 * x #2*TVday.AUC/30/30 = 0.1077018

#set.seed(143531)
set.seed(1432)
k = -0.07
Vt = V0 * exp(k*day)
n = length(Vt)
Vtt = Vt + rnorm(n, Vt, 0.5*day)
Vtt[Vtt<0] = 0
Vtt[9]=1
Vtt[10]=1
plot(Vtt~day)
TVday2 = data.frame(day,log(Vtt), log(Vtt[1]))
TVday2 = head(TVday2,-1)
names(TVday2) = c('day','logTV','logTV0')
TVday2.AUC = pracma::trapz(x=TVday2$day,y=TVday2$logTV) - pracma::trapz(x=TVday2$day,y=TVday2$logTV0)
TVday2.eg = function(x) log(100)  -0.08622278 * x #2*TVday.AUC/30/30 =  -0.08622278


pdf(file='Figure1a.pdf')
ggplot(TVday, aes(x=day, y=logTV)) +
  geom_point(color='steelblue3', size=5)+
  geom_line(aes(y = logTV), size=1.5, color='steelblue3') +
  geom_line(aes(y = logTV0), size=1.5) +
  geom_function(fun=TVday.eg, color='red', size=2) +
  geom_ribbon(data=TVday,
              aes(ymin=logTV0,ymax=logTV), fill="coral", alpha=0.3) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10),limits=c(0,30)) +
  #scale_y_continuous(breaks = seq(0, 8, by = 1), limits=c(0,8))+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10),limits=c(0,8))+
  geom_text(x=20, y=5.5, label="AUC=48.47", size=7,  color='black')+
  #geom_text(x=15, y=2, label=expression(nAUC == frac(AUC, 30^2)), size=9,  color='black')+
  geom_text(x=15, y=2, label=expression(nAUC == frac(2*AUC, 30^2)), size=9,  color='black')+
  geom_text(x=25, y=2, label="= 0.108", size=9,  color='black')+
  theme_pubr(base_size = 20)

ggplot(TVday2, aes(x=day, y=logTV)) +
  geom_point(color='steelblue3', size=5)+
  geom_line(aes(y = logTV), size=1.5, color='steelblue3') +
  geom_line(aes(y = logTV0), size=1.5) +
  geom_function(fun=TVday2.eg, color='red', size=2, xlim=c(0,27)) +
  geom_ribbon(data=TVday2,
              aes(ymin=logTV0,ymax=logTV), fill="lightgreen", alpha=0.3) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), limits=c(0,30)) +
  #scale_y_continuous(breaks = seq(0, 8, by = 1), limits=c(0,8))+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10),limits=c(0,8))+
  geom_text(x=20, y=4, label="AUC=-31.43", size=7, color='black')+
  geom_text(x=15, y=6, label=expression(nAUC == frac(2*AUC, 27^2)), size=9,  color='black')+
  geom_text(x=26, y=6, label="= -0.086", size=9,  color='black')+
  theme_pubr(base_size = 20)
dev.off()

t=35
tv=simu_TV2(cv_r=0.1,t=t,tgi_c=0.8,n = 10,vehicle_double_time = 6)
t_vec=round(seq(0,t,3.5)) #define vector of days to collect TV
if(t>max(t_vec)) t_vec=c(t_vec,t)
tv=tv %>% filter(Day %in% t_vec)
tv %>% ggplot() + aes(Day, TV, group = Mouse, color = Group) +
  geom_point() + geom_line() + theme_Publication() + facet_wrap(~Treatment) +
  ggsci::scale_color_jco() + theme(axis.title.x=element_text(face = "bold",size = rel(1.4)),axis.text=element_text(size=rel(1.6)),axis.ticks=element_line(size=rel(3)))+
  theme(legend.position = 'none')
ggsave("Fig1B_tumor_growth_curve_individual.png", width = 10,height = 8, dpi = 300)
auc_lst=get_mAUCr(tv)
auc_df=auc_lst$auc_mouse %>% group_by(Group) %>% summarise(nAUC=mean(AUC)) %>% left_join(tv %>% select(Group,Treatment) %>% unique())
tv_m=expand.grid(Day=seq(0,35,0.01),TV0=150,Treatment=levels(tv$Treatment)) %>% left_join(auc_df)
tv_m=tv_m %>% mutate(TV=TV0*exp(nAUC*Day))
tv_m %>% filter(TV<3000) %>% ggplot() + aes(Day, TV, color = Treatment)  + geom_line(size=1.2) +theme_Publication() +
  ggsci::scale_color_jco() +geom_vline(xintercept = 21,linetype=2,size=1.2)+
  theme(axis.title.x=element_text(face = "bold",size = rel(1.4)),axis.text=element_text(size=rel(1.6)),axis.ticks=element_line(size=rel(3)))+
  theme(legend.position = c(0.2,0.85),legend.title = element_blank(),legend.spacing.y = unit(0.1,'cm'),legend.key.size = unit(3,'line'),
        legend.text = element_text(size=rel(1.4),face='bold'),legend.direction = 'vertical')+guides(color=guide_legend(byrow = TRUE))
ggsave("Fig1C_tumor_growth_curve_nAUC.png", width = 10,height = 8, dpi = 300)
AUC_synergy(auc_lst,ci_type = 'bca',kw='Fig1D')
```

Data were downloaded from the link: https://static-content.springer.com/esm/art%3A10.1038%2Fs41586-022-04437-2/MediaObjects/41586_2022_4437_MOESM11_ESM.xlsx, which was related to Figure 8a of publication **Jaaks P, Coker E A, Vis D J, et al. Effective drug combinations in breast, colon and pancreatic cancer cells[J]. Nature, 2022, 603(7899): 166-173.**. CombPDX calculations were done from the webserver: https://licaih.shinyapps.io/CombPDX/, and the results were downloaded in Rdata format.

# Figure 2
```{r Figure2}
SW837 <- read_tv(system.file("extdata", "SW837.csv", package = "invivoSyn"))
plot_tumor_growth_curve(SW837,kw='Fig2A_SW837')
#load SW837 combPDX results
load('CombPDX results/SW837.Rdata')
bi_effect[2,4:5]=bi_effect[2,4:5]*10^15
level_order <- c("D7","D14","D21","D25","Global")
pval=bi_effect['Global','p.value']
bi_effect %>% filter(Days != 'D0') %>% mutate(Day=factor(Days,level=level_order)) %>% ggplot()+aes(Day,CI)+geom_point(color='blue',size=5)+
  geom_segment(aes(xend=Day,y=L95,yend=U95),size=1.5,color='black')+geom_line(size=1.3,color='gray',linetype=2)+
  geom_segment(aes(x=Day-0.2,xend=Day+0.2,y=L95,yend=L95),size=1.5,color='black')+
  geom_segment(aes(x=Day-0.2,xend=Day+0.2,y=U95,yend=U95),size=1.5,color='black')+xlab("Day")+ylab("Bliss CI(CombPDX)")+theme_Publication()+scale_x_discrete(limits=level_order)+
  theme(axis.title.x=element_text(face = "bold",size = rel(1.6)),axis.text=element_text(size=rel(1.5)),axis.ticks=element_line(size=rel(2.5)))+
  annotation_custom(grid::textGrob(label = paste0("gCI(CombPDX) pval=",pval),
                                    x = unit(0.7, "npc"), y = unit(0.95, "npc"),
                                    gp = grid::gpar(cex = 2)))
ggsave('Fig2B_SW837_CombPDX_CI.png',width=8,height=8,dpi=300)


AUC_lst <- get_mAUCr(SW837, ci = 0.95, ci_type = "bca")
bliss_synergy_AUC <- AUC_synergy(AUC_lst,ci_type = "bca",kw='Figure2C-2D_SW837_R')

SNU81 <- read_tv(system.file("extdata", "SNU-81.csv", package = "invivoSyn"))
plot_tumor_growth_curve(SNU81,kw='Fig2E_SNU81')
#load SNU-81 combPDX results
load('CombPDX results/SNU-81.Rdata')
bi_effect[2,4:5]=bi_effect[2,4:5]*10^15
level_order <- c("D8","D15","Global")
pval=bi_effect['Global','p.value']
bi_effect %>% filter(Days != 'D0') %>% mutate(Day=factor(Days,level=level_order)) %>% ggplot()+aes(Day,CI)+geom_point(color='blue',size=5)+
  geom_segment(aes(xend=Day,y=L95,yend=U95),size=1.5,color='black')+geom_line(size=1.3,color='gray',linetype=2)+
  geom_segment(aes(x=Day-0.2,xend=Day+0.2,y=L95,yend=L95),size=1.5,color='black')+ylim(c(-0.7,1.1))+
  geom_segment(aes(x=Day-0.2,xend=Day+0.2,y=U95,yend=U95),size=1.5,color='black')+xlab("Day")+ylab("Bliss CI(CombPDX)")+theme_Publication()+scale_x_discrete(limits=level_order)+
  theme(axis.title.x=element_text(face = "bold",size = rel(1.6)),axis.text=element_text(size=rel(1.5)),axis.ticks=element_line(size=rel(2.5)))+
  annotation_custom(grid::textGrob(label = paste0("gCI(CombPDX) pval=",pval),
                                   x = unit(0.7, "npc"), y = unit(0.95, "npc"),
                                   gp = grid::gpar(cex = 2)))
ggsave('Fig2F_SNU81_CombPDX_CI.png',width=8,height=8,dpi=300)
AUC_lst <- get_mAUCr(SNU81, ci = 0.95, ci_type = "bca")
bliss_synergy_AUC <- AUC_synergy(AUC_lst,ci_type = "bca",kw='Figure2G-2H_SNU-81')

LS1034 <- read_tv(system.file("extdata", "LS_1034.csv", package = "invivoSyn"))
plot_tumor_growth_curve(LS1034,kw='Fig2I_LS1034')
#load LS1034 combPDX results
load('CombPDX results/LS1034.Rdata')
bi_effect[2,4:5]=bi_effect[2,4:5]*10^15
bi_effect[2,4]=bi_effect[2,4]*10
level_order <- c("D3","D10","D17","Global")
pval=bi_effect['Global','p.value']
bi_effect %>% filter(Days != 'D0') %>% mutate(Day=factor(Days,level=level_order)) %>% ggplot()+aes(Day,CI)+geom_point(color='blue',size=5)+
  geom_segment(aes(xend=Day,y=L95,yend=U95),size=1.5,color='black')+geom_line(size=1.3,color='gray',linetype=2)+
  geom_segment(aes(x=Day-0.2,xend=Day+0.2,y=L95,yend=L95),size=1.5,color='black')+ylim(c(-0.7,2.0))+
  geom_segment(aes(x=Day-0.2,xend=Day+0.2,y=U95,yend=U95),size=1.5,color='black')+xlab("Day")+ylab("Bliss CI(CombPDX)")+theme_Publication()+scale_x_discrete(limits=level_order)+
  theme(axis.title.x=element_text(face = "bold",size = rel(1.6)),axis.text=element_text(size=rel(1.5)),axis.ticks=element_line(size=rel(2.5)))+
  annotation_custom(grid::textGrob(label = paste0("gCI(CombPDX) pval=",pval),
                                   x = unit(0.7, "npc"), y = unit(0.95, "npc"),
                                   gp = grid::gpar(cex = 2)))
ggsave('Fig2J_LS1034_CombPDX_CI.png',width=8,height=8,dpi=300)

AUC_lst <- get_mAUCr(LS1034, ci = 0.95, ci_type = "bca")
bliss_synergy_AUC <- AUC_synergy(AUC_lst,ci_type = 'bca',kw='Figure2L-2M_R_LS1034')
```

Depletion data is extracted from publication **Jin Y, An X, Mao B, et al. Different syngeneic tumors show distinctive intrinsic tumor-immunity and mechanisms of actions (MOA) of anti-PD-1 treatment[J]. Scientific Reports, 2022, 12(1): 3278.**, 

# Figure 3
```{r Figure3}
path='depletion_study/'
single_trt_info=read.csv(paste0(path,'single_trt.csv'),h=F)
colnames(single_trt_info)=c('Group','Treatment')
single_trt_info=single_trt_info %>% mutate(Treatment = gsub("\t","",Treatment),Group = gsub("0","",Group)) 
combo_trt_info=read.csv(paste0(path,'combo_trt.csv'),h=F)
colnames(combo_trt_info)=c('Group','Treatment')
trts=c('CD8','CD4','GM1','Liposomes','CD25')
grp_sels_comb=sapply(trts,function(tr) combo_trt_info %>% filter(grepl(tr,Treatment)) %>% pull(Group))
grp_sels_single=sapply(trts,function(tr) single_trt_info %>% filter(grepl(tr,Treatment) & !grepl('\\+',Treatment)) %>% pull(Group)) 
#combine auc for groups in two AUC list, vehicle group and aPD1 grows faster in single treatment experiment, while heterogeneity in higher in combo treatment experiment
combine_auc_lst=function(AUC_lst_c,AUC_lst_s,tr){
  auc_mouse_c=AUC_lst_c$auc_mouse
  auc_mouse_s=AUC_lst_s$auc_mouse
  auc_group_mean_s=AUC_lst_s$auc_mouse %>% group_by(Group) %>% summarise(m=mean(AUC)) %>% pull(m)
  auc_group_mean_c=AUC_lst_c$auc_mouse %>% group_by(Group) %>% summarise(m=mean(AUC)) %>% pull(m)
  cs_auc_ratio=auc_group_mean_c[1]/auc_group_mean_s[1]  #ratio of average nAUC between combo and single
  auc_mouse_c=AUC_lst_c$auc_mouse %>% filter(Group %in% c("Group 1","Group 2",grp_sels_comb[tr])) #%in% c("Group 1","Group 2",
  auc_mouse_c[auc_mouse_c$Group == grp_sels_comb[tr],'Group']="Group 4"
  max_mouse=max(auc_mouse_c$Mouse)
  #ajdust nAUC based on AUC ratio between vehicle
  auc_mouse_s=AUC_lst_s$auc_mouse %>% filter(Group %in%  c(grp_sels_single[tr])) %>% mutate(Mouse=Mouse+max_mouse,AUC=AUC*cs_auc_ratio)
  auc_mouse_s[auc_mouse_s$Group == grp_sels_single[tr],'Group']="Group 3"
  list(auc_mouse=rbind(auc_mouse_s,auc_mouse_c))
}
combine_auc_lst_special=function(AUC_lst_s){
  auc_mouse_s=AUC_lst_s$auc_mouse
  auc_mouse_s=auc_mouse_s %>% filter(Group %in% c("Group 1","Group 2","Group 7","Group 8"))
  auc_mouse_s[auc_mouse_s$Group == "Group 7",'Group']="Group 3"
  auc_mouse_s[auc_mouse_s$Group == "Group 8",'Group']="Group 4"
  list(auc_mouse=auc_mouse_s)
}
getSynergy=function(model,nboot=1000,trts=trts){
  dat1=read_tv(paste0(path,model,'_single.csv'))
  dat2=read_tv(paste0(path,model,'_combo.csv'))
  AUC_lst_s=get_mAUCr(dat1)
  AUC_lst_c=get_mAUCr(dat2)
  synergy_df=do.call(rbind,lapply(trts,function(trt)
  { 
      AUC_lst_cb=NULL
      if(model %in% c('MC38','Hepa16') & trt=='CD25')
        AUC_lst_cb=combine_auc_lst_special(AUC_lst_s)
      else
        AUC_lst_cb=combine_auc_lst(AUC_lst_c,AUC_lst_s,trt)
      AUC_synergy_cb=AUC_synergy(AUC_lst_cb,ci_type = "bca",boot_n = nboot,save=F,display = F)
      AUC_synergy_cb = AUC_synergy_cb %>% mutate(Treatment=trt)
      AUC_synergy_cb
    }))
  synergy_df=synergy_df %>% mutate(Model=model)
  list(Synergy_df=synergy_df,tv=list(single=dat1,combo=dat2),nAUC=list(single=AUC_lst_s,combo=AUC_lst_c))
}
#change GM1 to aNK, CD.. to aCD..
CT26_synergy=getSynergy('CT26',nboot=1000,trts=trts)
###EMT6, same experiment design; Hepa16, single treatment only have 6 groups;Hepa16 & MC38, anti-gr1(MDSC) instead of anti-CD25 (Treg)
EMT6_synergy=getSynergy('EMT6',nboot=1000,trts=trts)
#For MC38
single_trt_info=read.csv(paste0(path,'single_trt_MC38.csv'),h=F)
colnames(single_trt_info)=c('Group','Treatment')
single_trt_info=single_trt_info %>% mutate(Treatment = gsub("\t","",Treatment),Group = gsub("0","",Group)) 


combo_trt_info=read.csv(paste0(path,'combo_trt_MC38.csv'),h=F)
colnames(combo_trt_info)=c('Group','Treatment')
combo_trt_info=combo_trt_info %>% mutate(Treatment = gsub("\t","",Treatment),Group = gsub("0","",Group)) 


trts=c('CD8','CD4','NK','Liposomes','CD25')
grp_sels_comb=sapply(trts,function(tr) combo_trt_info %>% filter(grepl(tr,Treatment)) %>% pull(Group))
grp_sels_single=sapply(trts,function(tr) single_trt_info %>% filter(grepl(tr,Treatment) & !grepl('\\+',Treatment)) %>% pull(Group)) 
MC38_synergy=getSynergy('MC38',nboot=1000,trts=trts)

#For Hepa16
single_trt_info=read.csv(paste0(path,'single_trt_Hepa16.csv'),h=F)
colnames(single_trt_info)=c('Group','Treatment')
single_trt_info=single_trt_info %>% mutate(Treatment = gsub("\t","",Treatment),Group = gsub("0","",Group)) 


combo_trt_info=read.csv(paste0(path,'combo_trt_Hepa16.csv'),h=F)
colnames(combo_trt_info)=c('Group','Treatment')
combo_trt_info=combo_trt_info %>% mutate(Treatment = gsub("\t","",Treatment),Group = gsub("0","",Group)) 


trts=c('CD8','CD4','NK','Liposomes','CD25')
grp_sels_comb=sapply(trts,function(tr) combo_trt_info %>% filter(grepl(tr,Treatment)) %>% pull(Group))
grp_sels_single=sapply(trts,function(tr) single_trt_info %>% filter(grepl(tr,Treatment) & !grepl('\\+',Treatment)) %>% pull(Group)) 
Hepa16_synergy=getSynergy('Hepa16',nboot=1000,trts=trts)

all_synergy=rbind(CT26_synergy$Synergy_df,EMT6_synergy$Synergy_df,MC38_synergy$Synergy_df,Hepa16_synergy$Synergy_df)
all_synergy %>% count(Treatment)
all_synergy[all_synergy$Treatment %in% c('GM1','NK'),'Treatment']='aNK'
all_synergy=all_synergy %>% mutate(Treatment=gsub('CD','aCD',Treatment))

all_synergy %>% filter(Metric=='CI') %>% ggplot()+aes(Treatment,Value)+geom_point(color='blue',size=5)+geom_hline(yintercept = 1,lty=2,color='red')+
  geom_segment(aes(xend=Treatment,y=lb,yend=ub),size=1.5,color='black')+
  theme(axis.title.x=element_text(face = "bold",size = rel(1.4)),axis.text=element_text(size=rel(1.8)),axis.ticks=element_line(size=rel(8)))+
  facet_wrap(~Model,ncol=1,scales = 'free_y')+theme_Publication()+scale_y_log10()+ylab("Bliss CI(invivoSyn)")
ggsave("Fig3-CI_depletion.png",width=8,height=10,dpi=300)
```

We run the simulation regular tumor growth. Because it takes many hours to run simulation using a multi-core laptop, we saved the results of power calculation in a csv file.

# Figure 4
```{r}
#### Scenario A
kw='regular'
t=21
tv1=simu_TV(cv_r=0.1,vehicle_double_time = 7,tv0 = 100,n = 10,t=t)
t_vec=round(seq(0,t,3.5)) #define vector of days to collect TV
if(t>max(t_vec)) t_vec=c(t_vec,t)
tv1=tv1 %>% filter(Day %in% t_vec)
plot_tumor_growth_curve(tv1,kw='Fig4A')

#it will take many hours to run the simulations, therefore I saved the results in csv files
system.time(power_stats1 <- sim_power(
  type = c("CombPDX_CI", "AUC"),
  vehicle_double_time = 7,
  tv0=100,
  method = c("Bliss"),
  n_sim=100,
  n = 2:10,
  tgi_c = c(0.7,0.75,0.8)
))
write.csv(power_stats1,"power_stats_regular.csv",row.names = F)
power_stats1=read.csv('power_stats_regular_variable_tgr.csv')
power_stats1=power_stats1 %>% filter(vehicle_double_time==7) %>% mutate(level=case_when(tgi_c==0.70 ~ 'Nil',tgi_c==0.75 ~ 'Weak',tgi_c==0.8 ~ 'Strong')) %>%
  mutate(level=factor(level,levels=c("Nil","Weak","Strong")))
ggplot(power_stats1 %>% filter(type=='AUC')) +
  aes(n, Power) +
  #geom_point() +
  geom_smooth(n=3) +
  facet_wrap( ~ level) +
  theme_Publication() +
  scale_colour_Publication()+ylab('Empirical power')+
  theme(axis.text=element_text(size=rel(1.4)),axis.ticks=element_line(size=rel(2.5)),axis.title = element_text(size=rel(1.4)))
ggsave(paste0('Fig4B_',kw,"_power_curve_bliss.png"), width = 8,height = 6, dpi = 300)
ggsave(paste0('Fig4B_',kw,"_power_curve_bliss.pdf"), device = 'pdf',width = 8,height = 6, dpi = 300)


#vary tumor growth rate (Doubling time)
kw='regular_variable_tgr'
system.time(power_stats_r1 <- sim_power(
    type = c("CombPDX_CI", "AUC"),
    vehicle_double_time = c(3,5,7,10),
    tv0=100,
    method = c("Bliss"),
    n_sim=100,
    n = 2:10,
    tgi_c = c(0.7,0.75,0.8)
))
write.csv(power_stats_r1,"power_stats_regular_variable_tgr.csv",row.names = F)
power_stats_r1=read.csv('power_stats_regular_variable_tgr.csv')

power_stats_r1=power_stats_r1 %>% mutate(level=case_when(tgi_c==0.70 ~ 'Nil',tgi_c==0.75 ~ 'Weak',tgi_c==0.8 ~ 'Strong')) %>%
  mutate(level=factor(level,levels=c("Nil","Weak","Strong"))) %>% mutate(vehicle_double_time=factor(paste0("DT=",vehicle_double_time),
                                                                    levels=paste0("DT=",sort(unique(vehicle_double_time)))))
#the higher the tumor growth rate, the smaller the power (higher tumor growth rate -> higher variability of tumor volume)
ggplot(power_stats_r1 %>% dplyr::filter(type=='AUC')) +
    aes(n, Power, color = type, group = type) +
    geom_smooth(n=3) +
    #geom_line() +
    facet_grid(level~vehicle_double_time) +
    theme_Publication()+ylab('Empirical power')+
    scale_colour_Publication()+theme(legend.position = 'none')+
    theme(axis.text=element_text(size=rel(1.4)),axis.ticks=element_line(size=rel(2.5)),axis.title = element_text(size=rel(1.4)))
ggsave(paste0('Fig4C_',kw,"_power_curve_bliss.png"), width = 8,height = 6, dpi = 300)
ggsave(paste0('Fig4C_',kw,"_power_curve_bliss.pdf"), device = 'pdf',width = 10,height = 8, dpi = 300)


kw='regular_variable_euTV'
system.time(power_stats_r2 <- sim_power(
    type = c("CombPDX_CI", "AUC"),
    vehicle_double_time = 5,
    tv0=150,
    method = c("Bliss"),
    eu_tv=c(500,1000,1500,2000,2500),
    n_sim=100,
    n = 2:10,
    tgi_c = c(0.7,0.75,0.8)
))
power_stats_r2=read.csv('power_stats_regular_variable_euTV.csv')

#note: parse function can not parse character string starts with a number, and the parse function is not able to parse such character strings without further specifications. To solve that, we must explicitly surround the character string with quote
power_stats_r2=power_stats_r2 %>% mutate(level=case_when(tgi_c==0.70 ~ 'Nil',tgi_c==0.75 ~ 'Weak',tgi_c==0.8 ~ 'Strong')) %>%
  mutate(level=factor(level,levels=c("Nil","Weak","Strong"))) %>% mutate(eu_tv=factor(paste0("'",eu_tv," mm'^3"),
                                                                    levels=paste0("'",sort(unique(eu_tv))," mm'^3")))
ggplot(power_stats_r2 %>% dplyr::filter(type=='AUC')) +
    aes(n, Power, color = type, group = type) +
    geom_smooth(n=3) +
    #geom_line() +
    facet_grid(level~eu_tv,labeller =label_parsed)+ #use label_parsed here for superscript
    theme_Publication() +ylab('Empirical power')+
    scale_colour_Publication()+theme(legend.position = 'none')+
    theme(axis.text=element_text(size=rel(1.4)),axis.ticks=element_line(size=rel(2.5)),axis.title = element_text(size=rel(1.4)))
ggsave(paste0('Fig4D_',kw,"_power_curve_bliss.png"), width = 8,height = 6, dpi = 300)
ggsave(paste0('Fig4D_',kw,"_power_curve_bliss.pdf"), device = 'pdf',width = 10,height = 8, dpi = 300)
```

CombPDX calculations were done from the webserver: https://licaih.shinyapps.io/CombPDX/, and the results were downloaded in Rdata format.

# Figure 5
```{r}
CR1197 <- read_tv(system.file("extdata", "CR1197.csv", package = "invivoSyn"))
plot_tumor_growth_curve(CR1197,kw='Fig5A_CR1197')
load('CombPDX results/CR1197.Rdata')
bi_effect[2,4:5]=bi_effect[2,4:5]*10^15
level_order <- c("D3","D7","D10","D14","D17","Global")
pval=bi_effect['Global','p.value']
bi_effect %>% filter(Days != 'D0') %>% mutate(Day=factor(Days,level=level_order)) %>% ggplot()+aes(Day,CI)+geom_point(color='blue',size=5)+
  geom_segment(aes(xend=Day,y=L95,yend=U95),size=1.5,color='black')+geom_line(size=1.3,color='gray',linetype=2)+
  geom_segment(aes(x=Day-0.2,xend=Day+0.2,y=L95,yend=L95),size=1.5,color='black')+
  geom_segment(aes(x=Day-0.2,xend=Day+0.2,y=U95,yend=U95),size=1.5,color='black')+xlab("Day")+ylab("Bliss CI(CombPDX)")+theme_Publication()+scale_x_discrete(limits=level_order)+
  theme(axis.title.x=element_text(face = "bold",size = rel(1.6)),axis.text=element_text(size=rel(1.5)),axis.ticks=element_line(size=rel(2.5)))+
  annotation_custom(grid::textGrob(label = paste0("gCI(CombPDX) pval=",pval),
                                   x = unit(0.7, "npc"), y = unit(0.95, "npc"),
                                   gp = grid::gpar(cex = 2)))
ggsave('Fig5B_CR1197_CombPDX_CI.png',width=8,height=8,dpi=300)
AUC_lst <- get_mAUCr(CR1197, ci = 0.95, ci_type = "bca")
bliss_synergy_AUC <- AUC_synergy(AUC_lst,ci_type = 'bca',kw='Figure5C-5D_CR1197')

####Scenario B
t=50
kw='Delayed_kickin'
tv2=simu_TV(cv_r=0.05,vehicle_double_time = 5,tv0 = 100,tgi_c = 0.85,n = 10,t=t,kit_a = 10 ,kit_c = 10)
plot_tumor_growth_curve(tv2,y='logTV')
t_vec=round(seq(0,t,3.5)) #define vector of days to collect TV
if(t>max(t_vec)) t_vec=c(t_vec,t)
tv2=tv2 %>% filter(Day %in% t_vec)
plot_tumor_growth_curve(tv2,y='TV',kw='Fig5E')

system.time(power_stats3 <- sim_power(
  type = c("CombPDX_CI", "AUC"),
  vehicle_double_time = 5,
  tv0=100,
  method = c("Bliss"),
  n_sim=100,
  n = 2:10,
  tgi_c = c(0.85),
  t=c(30,40,50),
  kit_a=10,
  kit_c=10
))
write.csv(power_stats3,"power_stats_delayed_kickin_Bliss.csv",row.names = F)
power_stats3=read.csv('power_stats_delayed_kickin_Bliss.csv')
power_stats3=power_stats3 %>% mutate(t=paste0('T(obs)=',t, ' days'))
ggplot(power_stats3 %>% filter(type=='AUC')) +
  aes(n, Power) +
  #geom_point() +
  geom_smooth(n=3) +
  facet_wrap(~t) +
  theme_Publication() +ylab('Empirical power')+
  scale_colour_Publication()+scale_y_continuous(breaks = seq(0,1,0.2))+
  theme(axis.text=element_text(size=rel(1.4)),axis.ticks=element_line(size=rel(2.5)),axis.title = element_text(size=rel(1.4)))
ggsave(paste0('Fig5F_',kw,"_power_curve_bliss.png"), width = 8,height = 6, dpi = 300)

```


# Figure 6

```{r}
###Scenario C
t=50
kw='Intrinsic_resistance'
tv3=simu_TV(cv_r=0.05,vehicle_double_time = 7,tv0 = 200,tgi_c = 0.8,n = 10,t=t,rp_a = 0.1,rp_b=0.1,rp_c=0.1)
t_vec=round(seq(0,t,3.5)) #define vector of days to collect TV
if(t>max(t_vec)) t_vec=c(t_vec,t)
tv3=tv3 %>% filter(Day %in% t_vec)
plot_tumor_growth_curve(tv3,y='TV',kw='Fig6A')

system.time(power_stats5 <- sim_power(
  type = c("CombPDX_CI", "AUC"),
  cv_r=c(0.1),
  vehicle_double_time = 5,
  tv0=100,
  method = c("Bliss"),
  n_sim=100,
  n = 2:10,
  tgi_c = c(0.75,0.8,0.9),
  t=c(30,40,50),
  rp_a=0.1,
  rp_b=0.1,
  rp_c=0.1
))
write.csv(power_stats5,"power_stats_intrinsic_resistance_Bliss.csv",row.names = F)
power_stats5=read.csv('power_stats_intrinsic_resistance_Bliss.csv')
power_stats5=power_stats5 %>% mutate(t=paste0('T(obs)=',t, ' days'))
ggplot(power_stats5 %>% filter(type=='AUC')) +
  aes(n, Power) +
  #geom_point() +
  geom_smooth(n=3) +
  facet_wrap(~t) +
  theme_Publication() +ylab('Empirical power')+
  scale_colour_Publication()+scale_y_continuous(breaks = seq(0,1,0.2))+
  theme(axis.text=element_text(size=rel(1.4)),axis.ticks=element_line(size=rel(2.5)),axis.title = element_text(size=rel(1.4)))
ggsave(paste0('Fig6B_',kw,"_power_curve_bliss.png"), width = 8,height = 6, dpi = 300)

###Scenario D
t=50
kw='Induced_resistance'
tv4=simu_TV(cv_r=0.05,vehicle_double_time = 7,tv0 = 200,tgi_c = 0.8,n = 10,t=t,rp_a = 0.1,rp_c=0.1,irt_a = 10, irt_c = 10)

t_vec=round(seq(0,t,3.5)) #define vector of days to collect TV
if(t>max(t_vec)) t_vec=c(t_vec,t)
tv4=tv4 %>% filter(Day %in% t_vec)
plot_tumor_growth_curve(tv4,y='TV',kw='Fig6C')

system.time(power_stats6 <- sim_power(
  type = c("CombPDX_CI", "AUC"),
  cv_r=c(0.05),
  vehicle_double_time = 7,
  tv0=200,
  method = c("Bliss"),
  n_sim=100,
  n = 2:10,
  tgi_c = c(0.75,0.8),
  t=c(30,40,50,60),
  rp_a=0.1,
  rp_c=0.1,
  irt_a=10,
  irt_c=10
))
write.csv(power_stats6,'power_stats_induced_resistance_Bliss.csv',row.names = F)
power_stats6=read.csv('power_stats_induced_resistance_Bliss.csv')
power_stats6=power_stats6 %>% filter(type=='AUC' & t != 50 & tgi_c==0.8)
power_stats6=power_stats6 %>% mutate(t=paste0('T(obs)=',t, ' days'))
ggplot(power_stats6 %>% filter(type=='AUC')) +
  aes(n, Power) +
  #geom_point() +
  geom_smooth(n=3) +
  facet_wrap(~t) +
  theme_Publication() +ylab('Empirical power')+
  scale_colour_Publication()+scale_y_continuous(breaks = seq(0,1,0.2))+
  theme(axis.text=element_text(size=rel(1.4)),axis.ticks=element_line(size=rel(2.5)),axis.title = element_text(size=rel(1.4)))
ggsave(paste0('Fig6D_',kw,"_power_curve_bliss.png"), width = 8,height = 6, dpi = 300)
```

# Figure 7
```{r}
###Ex vivo, survival (time vs survival, different dose level)
dose_level=1/10^seq(0,6,1)
pals=ggpubr::get_palette("jco",4)
t_sel=c(24,48,72,96)
t=0:96
df=expand.grid(t=t,dose=dose_level)
df=df %>% filter(dose!=1E-6) %>% mutate(k_t=(6+log10(dose))*0.01) %>% mutate(Survival=exp(-k_t*t)) %>% mutate(dose=factor(paste0(dose,' mg/L'),levels=paste0(dose_level,' mg/L')))
df %>% ggplot()+aes(t,Survival,group=dose,color=dose)+geom_line(lwd=1.5)+geom_vline(xintercept = 72,lty=2,color=pals[4],lwd=2)+
  theme_Publication()+ggsci::scale_color_jco()+theme(axis.title.x=element_text(face = "bold",size = rel(1.4)),axis.text=element_text(size=rel(1.6)),axis.ticks=element_line(size=rel(3)))+
  theme(legend.position = c(0.85,0.8),legend.title = element_blank(),legend.spacing.y = unit(0.2,'cm'),legend.key.size = unit(2,'line'),
        legend.text = element_text(size=rel(1.4),face='bold'),legend.direction = 'vertical')+guides(color=guide_legend(byrow = TRUE))+
  scale_x_continuous(breaks=t_sel)+xlab('Time(hr)')+ylab("Relative Survival")
ggsave("Fig7A_survival_vs_time_exvivo.png", width = 8,height = 8, dpi = 300)

#dose_level <- seq(0, 8, 0.1)
dose_level=1/10^seq(0,8,0.01)
Treatment <- c("Drug A", "Drug B", "Drug A+B")
drug_info <- expand.grid(Drug = Treatment, Dose = dose_level)
drug_info <- drug_info %>%
  mutate(m = ifelse(Drug == "Drug A", 1E-3,
                    ifelse(Drug == "Drug B", 1E-2, 1E-4)))
b=2
drug_info=drug_info %>% mutate(Survival=1/(1+exp(b*(log(Dose)-log(m)))))
drug_info=drug_info %>% mutate(Drug=factor(Drug,levels=c("Vehicle","Drug A","Drug B","Drug A+B")))

drug_info %>% ggplot()+aes(Dose,Survival,group=Drug,color=Drug)+geom_line(lwd=1.5)+
  theme_Publication()+scale_color_manual(values=pals[-1])+theme(axis.title.x=element_text(face = "bold",size = rel(1.4)),axis.text=element_text(size=rel(1.6)),axis.ticks=element_line(size=rel(3)))+
  theme(legend.position = c(0.85,0.85),legend.title = element_blank(),legend.spacing.y = unit(0.2,'cm'),legend.key.size = unit(2,'line'),
        legend.text = element_text(size=rel(1.4),face='bold'),legend.direction = 'vertical')+guides(color=guide_legend(byrow = TRUE))+
  xlab('Dose(mg/L)')+ylab("Relative Survival")+ylim(c(0,1))+scale_x_log10(breaks = 1/10^seq(0,8,2))
ggsave("Fig7B_curve_shift_exvivo.png", width = 8,height = 8, dpi = 300)

ci_df=data.frame('FA'=1:100) %>% mutate(CI=0.9-FA/200+rnorm(100,0,0.15))
ci_df %>% ggplot()+aes(FA,CI)+geom_smooth(lwd=1.5)+
  theme_Publication()+theme(axis.title.x=element_text(face = "bold",size = rel(1.4)),axis.text=element_text(size=rel(1.6)),axis.ticks=element_line(size=rel(3)))+
  xlab('FA(%)')+ylab("CI")+ylim(c(0,1))+geom_hline(yintercept = 1,lty=2,lwd=2,color=pals[4])
ggsave("Fig7C_FA_CI_exvivo.png", width = 8,height = 8, dpi = 300)

dose_level=c(0,2,5,10,20,35,50)
t_vec=0:30
tgi_a=c(0,0.1,0.3,0.6,0.7,0.8,0.9)
k_v=log(2)/7 #calcul-1ate growth rate for vehicle group
tv=expand.grid(dose=dose_level,t=t_vec)
tv=tv %>% mutate(TV0=100) %>% left_join(data.frame(dose=dose_level,tgi=tgi_a))
tv=tv %>% mutate(k=k_v+log(1-tgi)/21) %>% mutate(TV=TV0*exp(k*t))
tv=tv %>% mutate(dose=factor(paste0(dose,' mg/kg'),levels=paste0(dose_level,' mg/kg')))
tv %>% filter(TV<3000) %>% ggplot() + aes(t, TV, color = dose)  + geom_line(size=1.2) +theme_Publication() +
  ggsci::scale_color_jco() +
  theme(axis.title.x=element_text(face = "bold",size = rel(1.4)),axis.text=element_text(size=rel(1.6)),axis.ticks=element_line(size=rel(3)))+
  theme(legend.position = c(0.15,0.8),legend.title = element_blank(),legend.spacing.y = unit(0.1,'cm'),legend.key.size = unit(2,'line'),
        legend.text = element_text(size=rel(1.4),face='bold'),legend.direction = 'vertical')+guides(color=guide_legend(byrow = TRUE))+xlab('Time(Day)')
ggsave("Fig7D_tumor_growth_curve_invivo.png", width = 8,height = 8, dpi = 300)

t=30
tv_combo=simu_TV2(cv_r=0.1,t=t,tgi_c=0.7,n = 10,vehicle_double_time = 6)
t_vec=round(seq(0,t,3.5)) #define vector of days to collect TV
if(t>max(t_vec)) t_vec=c(t_vec,t)
tv_combo=tv_combo %>% filter(Day %in% t_vec)
auc_lst=get_mAUCr(tv_combo)
auc_df=auc_lst$auc_mouse %>% group_by(Group) %>% summarise(nAUC=mean(AUC)) %>% left_join(tv_combo %>% select(Group,Treatment) %>% unique())
tv_m2=expand.grid(Day=seq(0,35,0.01),TV0=150,Treatment=levels(tv_combo$Treatment)) %>% left_join(auc_df)
tv_m2=tv_m2 %>% mutate(TV=TV0*exp(nAUC*Day))
tv_m2 %>% filter(TV<3000) %>% ggplot() + aes(Day, TV, color = Treatment)  + geom_line(size=1.2) +theme_Publication() +
  ggsci::scale_color_jco() +geom_vline(xintercept = 21,linetype=2,size=1.2)+
  theme(axis.title.x=element_text(face = "bold",size = rel(1.4)),axis.text=element_text(size=rel(1.6)),axis.ticks=element_line(size=rel(3)))+
  theme(legend.position = c(0.3,0.85),legend.title = element_blank(),legend.spacing.y = unit(0.1,'cm'),legend.key.size = unit(2,'line'),
        legend.text = element_text(size=rel(1.4),face='bold'),legend.direction = 'vertical')+guides(color=guide_legend(byrow = TRUE))+xlab("Time(Day)")
ggsave("Fig7E_tumor_growth_curve_nAUC.png", width = 8,height = 8, dpi = 300)


t_vec=seq(0,30,0.1)
ci_df=data.frame(t_vec) %>% mutate(CI=1-0.4*t_vec/30+rnorm(length(t_vec),0,0.15))
ci_df %>% ggplot()+aes(t_vec,CI)+geom_smooth(lwd=1.5)+
  theme_Publication()+theme(axis.title.x=element_text(face = "bold",size = rel(1.4)),axis.text=element_text(size=rel(1.6)),axis.ticks=element_line(size=rel(3)))+
  xlab('Time(Day)')+ylab("CI")+geom_hline(yintercept = 1,lty=2,lwd=2,color=pals[4])
ggsave("Fig7F_Day_CI_invivo.png", width = 8,height = 8, dpi = 300)

```


# Reference

1.  **Guo S, Jiang X, Mao B, et al.** The design, analysis and application of mouse clinical trials in oncology drug development[J]. *BMC cancer*, 2019, 19(1): 1-14.
2.  **Demidenko E, Miller T W.** Statistical determination of synergy based on Bliss definition of drugs independence[J]. *PLoS One*, 2019, 14(11): e0224137.
3.  **Huang L, Wang J, Fang B, et al.** CombPDX: a unified statistical framework for evaluating drug synergism in patient-derived xenografts[J]. *Scientific reports*, 2022, 12(1): 1-10.
